---
title: "p8105_hw2_yw4199"
author: "Yaduo Wang"
date: "2023-10-01"
output: github_document
---
Problem 2

```{r}
library(tidyverse)
library(readxl)
```

```{r}
trash_data = read_excel("data_files/Trash_Collection_Data.xlsx", sheet = 1, skip = 1)
#specify the sheet in the Excel file and to omit non-data entries
trash_data = janitor::clean_names(trash_data)
#use reasonable variable names

view(trash_data)

Mr.trash = 
  trash_data |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> 
#omit rows that do not include dumpster-specific data
  mutate(homes_powered = weight_tons*500/30)
#Update the data to include a new homes_powered variable based on this calculation.

```

```{r}
Pro_data = read_excel("data_files/Trash_Collection_Data.xlsx", sheet = 2, skip = 1)
#specify the sheet in the Excel file and to omit non-data entries
Pro_data = janitor::clean_names(Pro_data)
#use reasonable variable names
view(Pro_data)

Pro.trash = 
  Pro_data |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> 
#omit rows that do not include dumpster-specific data
  mutate(homes_powered = weight_tons*500/30)
#Update the data to include a new homes_powered variable based on this calculation.
view(Pro.trash)
```

```{r}
Gwy_data = read_excel("data_files/Trash_Collection_Data.xlsx", sheet = 4, skip = 1)
#specify the sheet in the Excel file and to omit non-data entries
Gwy_data = janitor::clean_names(Gwy_data)
#use reasonable variable names
view(Gwy_data)

Gwy.trash = 
  Gwy_data |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> 
#omit rows that do not include dumpster-specific data
  mutate(homes_powered = weight_tons*500/30)
#Update the data to include a new homes_powered variable based on this calculation.
view(Gwy.trash)
```

Combine all three datasets.
```{r}
#add an additional variable to all datasets before combining
Mr.trash = 
  Mr.trash |> 
  mutate(name_sheet = "Mr.Trash Wheel")
Pro.trash = 
  Pro.trash |> 
  mutate(name_sheet = "Professor Trash Wheel") 
Gwy.trash = 
  Gwy.trash |> 
  mutate(name_sheet = "Gwynnda Trash Wheel") 

#keep the variable type constant
Mr.trash$year = as.numeric(Mr.trash$year)
Pro.trash$year = as.numeric(Pro.trash$year)
Gwy.trash$year = as.numeric(Gwy.trash$year)
```

```{r}
#combine all three datasets
three_tidy = 
  bind_rows(Mr.trash, Pro.trash, Gwy.trash) |>
  janitor::clean_names()
view(three_tidy)
```

Description of the data
```{r}
sum_weight_Professor = filter(three_tidy, name_sheet == "Professor Trash Wheel")
sum_gwy = filter(three_tidy, 
                 month == "July", year == 2021, name_sheet =="Gwynnda Trash Wheel")
```


This new dataset contains `r nrow(three_tidy)` rows, and `r ncol(three_tidy)` columns. The columns in this dataset include: `r names(three_tidy)`. The total weight of trash collected by Professor Trash Wheel is `r sum(sum_weight_Professor$weight_tons)`. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(sum_gwy$cigarette_butts)`.


Problem 3
```{r}

```

